Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-06-30 doc/RMD/mlg-mcg.Rmd
update barplot figure to fix #2

```{r barplots, fig.width = 5, fig.height = 15}
mcg_graph <- mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), count) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  coord_flip() +
  theme(panel.grid.major.y = element_blank()) +
  theme(legend.position = c(.745, .85)) +
  # theme(legend.box = "horizontal") +
  theme(aspect.ratio = 3) +
  labs(list(
    alpha = "MLG Evenness"
  ))

uneven <- mcgs %>% 
  filter(MCG == 5) %>% 
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
  geom_col(color = "grey62", fill = "grey72", width = 0.5) + 
  theme_void() +
  ggtitle("Evenness = 0.48") +
  theme(aspect.ratio = 1) +
  theme(title = element_text(hjust = 0.5, 
                             size = rel(1.25),
                             family = "Helvetica",
                             color = "grey30"))

even <- mcgs %>% 
  filter(MCG == 53) %>% 
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
  geom_col(color = "grey20", fill = "grey20", width = 0.75) + 
  theme_void() +
  ggtitle("Evenness = 1") +
  theme(aspect.ratio = 1) +
  theme(title = element_text(hjust = 0.5, 
                             size = rel(1.25),
                             family = "Helvetica",
                             color = "grey30"))

vp1 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.67, just = "right")
vp2 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.555, just = "right")

pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 12)
print(mcg_graph)
print(uneven, vp = vp1)
print(even, vp = vp2)
dev.off()

print(mcg_graph)
print(uneven, vp = vp1)
print(even, vp = vp2)
```

@@

```{r}
make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}

>>>>

```{r barplots, fig.width = 5, fig.height = 10}
mcg_graph <- mcgs %>% 
  filter(MLGs > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), count) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  coord_flip() +
  theme(panel.grid.major.y = element_blank()) +
  theme(legend.box.just = "left") +
  theme(legend.box.margin = unit(c(0, 0.55, 11, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = NA, color = "black")) +
  # theme(legend.box = "horizontal") +
  theme(aspect.ratio = 2) +
  labs(list(
    alpha = "MLG\nEvenness"
  ))

even <- mcgs %>% 
  filter(MCG %in% c(53, 9)) %>% 
  mutate(MCG = ifelse(MCG == 53, "Evenness = 1.00", "Evenness = 0.55")) %>%
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq, alpha = Evenness)) + 
  geom_col(color = "grey30", width = 0.75) + 
  theme_void() +
  theme(strip.text = element_text(family = "Helvetica", color = "grey30", face = "bold")) +
  theme(aspect.ratio = 0.66) +
  theme(panel.spacing = unit(0, "line")) +
  theme(legend.position = "none") +
  # theme(plot.background = element_rect(fill = NA, colour = "black")) +
  facet_wrap(~MCG, ncol = 1, scale = "free")

vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 0.795, y = 0.64, just = c("center", "top"))
mcg_graph <- mcg_graph + theme(legend.position = c(.789, .585))

pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 10)
print(mcg_graph)
print(even, vp = vp1)
dev.off()

print(mcg_graph)
print(even, vp = vp1)
```

@@

```{r}

make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}
